Security Properties for Stack Safety

### 2021-05-17 short/main.tex
Polish

Our threat model is very strong, allowing the attacker (i.e., the callee) to
execute arbitrary code, including code that might attempt to smash
the stack to disrupt the program's control flow.
Our stack safety properties demand that, even in the presence of such
attackers, confidentiality and integrity
still apply to data in suspended stack frames until the callee returns.
The question then arises:
what constitutes a return in this setting? It is not obvious, since our
setting is a concrete, low-level ISA where the same opcode might sometimes
be used to return from a call and sometimes for other purposes.
Our solution is to associate with each call a predicate on states,
its ``return target,''
which defines a correct return: the program counter
must be at the next instruction, and the stack pointer restored to its original
value. So the callee may
execute instructions outside of its normal footprint, perhaps due to
smashing the stack, but
the caller is protected until the first state that satisfies its return target.

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}---a global requirement that callees
must always return
to their immediate caller, if they return at all. This property is orthogonal to our
data-protection properties. It remains meaningful in our strong attacker model,
to support reasoning about the ordering of valid returns.

@@

We demonstrate our model's flexibility by extending it to different settings.
First, we can support passing arguments on the stack, allowing some of the state of
the caller to be shared with the callee (and potentially with a sub-callee,
and so on).  We can also describe a simple coroutine model, with one
statically bounded stack per coroutine, each protected from the others.

In ongoing work, we investigate whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we aim to validate
the ``uninitialized capability model'' of \citet{Georges+21} using
our testing framework. Out of several Cheri-based
stack protection schemes, their treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques can support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.

>>>>

Our threat model is very strong, allowing the attacker (i.e., the callee) to
execute arbitrary code, including attempting to smash
the stack to disrupt the program's control flow.
Our stack safety properties demand that, even in the presence of such
attackers, confidentiality and integrity
still apply to data in suspended stack frames until the callee returns.
The question then arises:
what constitutes a return in this setting? It is not obvious, since
our machine model is a low-level ISA where the same opcode might
sometimes
be used to return from a call and sometimes for other purposes.
Our solution is to associate, with each dynamic call in an execution trace, a
``return target'' that records the address of the instruction following the
call and the stack pointer when the call was made. Thus, so the callee may
execute instructions outside of its normal footprint, perhaps due to
smashing the stack, but
the caller is protected until the PC and SP simultaneously match its return
target.

We can also, optionally, include a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}---a global requirement that callees
must always return
to their immediate caller, if they return at all. This property is orthogonal to our
data-protection properties. It remains meaningful in our strong attacker model,
to support reasoning about the ordering of valid returns.

@@

We demonstrate our model's flexibility by extending it in two significant ways.
First, we can support passing arguments on the stack, allowing some of the state of
the caller to be shared with the callee (and potentially with a sub-callee,
and so on).  We can also describe a simple coroutine model, with one
statically bounded stack per coroutine, each protected from the others.

In ongoing work, we are investigating whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we aim to validate
the ``uninitialized capability model'' of \citet{Georges+21} using
our testing framework. Out of several Cheri-based
stack protection schemes, their treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques can support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.
